\textsc{Rils}-\textsc{Rols} : Robust Symbolic Regression via Iterated Local Search and Ordinary Least Squares

### 2022-11-08 paper_resources/doc/paper.tex
Abstract additinally revised

	\author[1]{Aleksandar Kartelj}
	\author[2]{Marko Djukanovi\'c}
	\address[1]{$kartelj@matf.bg.ac.rs$, \\  Faculty of Mathematics, University of Belgrade, Serbia}
	\address[2]{$ marko.djukanovic@pmf.unibl.org$,\\   Faculty of Natural Sciences and Mathematics, University of Banja Luka, Bosnia and Herzegovina}
	\begin{abstract}
		In this paper, we solve the well-known, prominent problem of Symbolic regression, intensively studied within the last two decades. This problem has a wide range of applications;  for example, the obtained SR models could serve theoretically, providing insight towards establishing physical theory behind, thus helping establish reasonable hypotheses.  Practical applications include revealing complex ecological dynamics in ecology, discovering mutations effects on protein stability in biology, modelling analytic representations of the exciton binding energy in energy science, and finding models for band gaps of NaCl-type compounds
		in material science, just to name a few. We propose a new meta-heuristic-based approach, called \textsc{Rils}-\textsc{Rols} algorithm to solve the problem of Symbolic regression. There are several key concepts utilized within this algorithm: ($i$) Iterated local search scheme is used as the backbone; ($ii$) ordinary least square method is used within the local search procedure to make it efficient and to fit the coefficients in models (solutions);  ($iii$) a novel fitness function is utilized which combines an accuracy score, error score, and the size of the model into one, more compact score. The experiments are conducted on the two ground-truth benchmark sets, \textsc{Feynman} and \textsc{Strogatz} comparing \textsc{Rils}-\textsc{Rols} to 14 other competitors from the literature considering different levels of noise in the input data.  Our method obtains new best exact percentages on both benchmark sets compared to the literature approaches. Moreover, a new unbiased randomly generated set of problem instances are introduced where the scalability and robustness of our method are thoroughly demonstrated in terms of various instance properties. 
	\end{abstract}
	\maketitle

>>>>

	\author[1]{Aleksandar Kartelj}
	\author[2]{Marko Djukanovi\'c}
	\address[1]{$kartelj@matf.bg.ac.rs$, \\  Faculty of Mathematics, University of Belgrade, Serbia}
	\address[2]{$ marko.djukanovic@pmf.unibl.org$,\\   Faculty of Natural Sciences and Mathematics, University of Banja Luka, Bosnia and Herzegovina}
	\begin{abstract}
		In this paper, we solve the well-known, prominent problem of Symbolic regression, intensively studied within the last two decades. This problem has a wide range of applications;  for example, the obtained SR models could serve theoretically, providing insight towards establishing physical theory behind, thus helping establish reasonable hypotheses.  Practical applications include revealing complex ecological dynamics in ecology, discovering mutations effects on protein stability in biology, modelling analytic representations of the exciton binding energy in energy science, and finding models for band gaps of NaCl-type compounds
		in material science, just to name a few. We propose a new meta-heuristic-based approach, called \textsc{Rils}-\textsc{Rols} algorithm to solve the problem of Symbolic regression. There are several key concepts utilized within this algorithm: ($i$) Iterated local search scheme is used as the backbone; ($ii$) ordinary least square method is used within the local search procedure to make it efficient and to   determine the best--fitting   coefficients in considered models (solutions);  ($iii$) a novel fitness function is utilized which combines an accuracy score, error score, and the weighted size of the model into one, more compact score. The experiments are conducted on the two ground-truth benchmark sets, \textsc{Feynman} and \textsc{Strogatz} comparing \textsc{Rils}-\textsc{Rols} algorithm to 14 other competitors from the literature considering different levels of noise in the input data.  Our method obtains new best exact percentages on both benchmark sets compared to other literature approaches. Moreover, a new unbiased randomly generated set of problem instances are introduced where the scalability of our method is thoroughly demonstrated in terms of various instance properties. 
	\end{abstract}
	\maketitle
